Tool-driven Revolutions in Archaeological Science

### 2018-09-10 analysis/paper/jcaa-paper.Rmd
rewrote intro to be a bit more precise in what we do in the paper

In this paper we investigate recent changes in the way computational archaeology is being done. Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discurses[@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science are either ideas-driven or tool-driven [@Galison_1997; @Dyson_2000]. In this light we explore how many scientific disciplines have started to embrace methods to improve the reproducibility of their research and whether this change is part of a paradigm-shift. We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages and review bibliometric evidence for this change in the sciences generally and in archaeology in particular. We evaluate how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and test if the adoption of these new tools may be considered a revolution or some other process.

[need some more in here]

@@

Starting with the discussion of Kuhn's influential work on paradigm-shifts in science [citation needed], we offer a second explanation on how pardigm-shifts develop in science with the help of Galison [-@Galison_1997], who focuses on the material culture of a science and its tools. A change in the paradigms of a discipline may also only occur, if tools to implement this change are available. The change we are aiming for is the move to a reproducible computational archaeology, that adheres to the scientific method. The tool which we have been discussing here is R, as an example for scientific scripting languages.

>>>>

In this paper we investigate recent developments which facilitate computational archaeology to return to principles of the scientific method and whether this might be the beginning of a paradigm-shift. Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discourses [@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science might not just be ideas-driven but also tool-driven [@Galison_1997; @Dyson_2000]. In this light we explore how Ecology and Archaeology have started to embrace methods which improve the reproducibility of their research. We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of code. A review of bibliometric evidence showcases a strong development towards the use of R for Ecology and hints at the beginning of a similar development in archaeology. We thus evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be the sign of a tool-driven paradigm-change in archaeology. Acknowledging this is not a simple process, we offer a R-based tool to facilitates the task of creating a compendium which enables other researchers to reproduce the published results.

Code and text for this paper are available at: https://github.com/benmarwick/caa2018-Tubingen

@@

Starting with the discussion of Kuhn's influential work on paradigm-shifts in science [@Kuhn_1962], we offer a second explanation on how pardigm-shifts develop in science with the help of Galison [-@Galison_1997], who focuses on the material culture of a science and its tools. A change in the paradigms of a discipline may also only occur, if tools to implement this change are available. The change we are aiming for is the move to a reproducible computational archaeology, that adheres to the scientific method. The tool which we have been discussing here is R, as an example for scientific scripting languages.
